Analysis of transit delays in the Toronto Transit Commission: 2024

### 2024-09-26 paper/paper.qmd
Update paper.qmd

subset_data <- data %>%
  select(Date, Time, Day, Incident, Min.Delay, Min.Gap) %>% 
  head(10)

@@

subset_data <- data %>%
  select(Date, Time, Day, Code, Min.Delay, Min.Gap) %>% 
  head(10)

@@

1. Date: The dataset shows the date when the delay occurred.
2. Time: The dataset contains information on the time the delay occurred.
3. Day: The dataset indicates the day of the week when the delay occurred.
4. Incident: The dataset displays the reason for the delay, such as a general delay, emergency services, or security issues. Incident is displayed in code for Table 3.
5. Min.Delay and Min.Gap: The dataset likely represents 'Min Delay' and 'Min Gap,' both measured in minutes. 'Min Delay' refers to the delay in the schedule for the next bus, while 'Min Gap' indicates the scheduled time, between the bus ahead and the following bus.

@@

# Create the plot with abbreviated day labels
ggplot(combined_data, aes(x = Day, y = Count, color = vehicle, group = vehicle)) +
  geom_line(size = 1) +
  geom_point() +
  labs(title = "Number of delays caused in day, by Vehicle",
       x = "Day of the Week",
       y = "Count of Delays",
       color = "Vehicle") +
  theme_minimal() +
  scale_x_discrete(limits = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))
```
Figure 3:The trend of delays based on the day of the week

@@

This information offers valuable insight for the TTC to prepare alternative solutions for delays. For instance, they could deploy more buses or trains on days with higher delays compared to days with fewer delays.

>>>>

subset_data <- data %>%
  select(Date, Time, Day, Incident, Min_Delay, Min_Gap) %>% 
  head(10)

@@

subset_data <- data %>%
  select(Date, Time, Day, Code, Min_Delay, Min_Gap) %>% 
  head(10)

@@

1. Date: The dataset shows the date when the delay occurred.
2. Time: The dataset contains information on the time the delay occurred.
3. Day: The dataset indicates the day of the week when the delay occurred.
4. Incident: The dataset displays the reason for the delay, such as a general delay, emergency services, or security issues. Incident is displayed in code for Table 3.
5. Min_Delay and Min_Gap: The dataset likely represents 'Min Delay' and 'Min Gap,' both measured in minutes. 'Min Delay' refers to the delay in the schedule for the next bus, while 'Min Gap' indicates the scheduled time, between the bus ahead and the following bus.

@@

# Create the plot with abbreviated day labels
ggplot(combined_data, aes(x = Day, y = Count, color = vehicle, group = vehicle)) +
  geom_line(size = 1) +
  geom_point() +
  labs(title = "Number of delays caused in day, by Vehicle",
       x = "Day of the Week",
       y = "Count of Delays",
       color = "Vehicle") +
  theme_minimal() +
  scale_x_discrete(limits = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))
```
Figure 3: The trend of delays based on the day of the week

@@

This information provides valuable insights for the TTC to prepare alternative solutions for delays. For instance, they could deploy more buses or trains on days with higher delays compared to those with fewer delays. In this case, the TTC could operate more buses on Tuesdays, as an upward trend in delays is observed from Monday to Tuesday. Conversely, the TTC could reduce the number of buses on weekends, as there is a downward trend in delays during that time.

# Summarize total number of delays and average delay time for each vehicle
bus_summary <- bus_data %>%
  summarise(Total_Delays = n(), Avg_Delay_Time = mean(`Min_Delay`, na.rm = TRUE))

streetcar_summary <- streetcar_data %>%
  summarise(Total_Delays = n(), Avg_Delay_Time = mean(`Min_Delay`, na.rm = TRUE))

subway_summary <- subway_data %>%
  summarise(Total_Delays = n(), Avg_Delay_Time = mean(`Min_Delay`, na.rm = TRUE))

# Combine the summaries into one data frame
summary_data <- data.frame(
  Vehicle = c("Bus", "Streetcar", "Subway"),
  Total_Delays = c(bus_summary$Total_Delays, streetcar_summary$Total_Delays, subway_summary$Total_Delays),
  Avg_Delay_Time = c(bus_summary$Avg_Delay_Time, streetcar_summary$Avg_Delay_Time, subway_summary$Avg_Delay_Time)
)

# Plot 1: Total Number of Delays per Vehicle
ggplot(summary_data, aes(x = Vehicle, y = Total_Delays, fill = Vehicle)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Number of Delays per Vehicle", x = "Vehicle", y = "Total Delays") +
  theme_minimal()

# Plot 2: Average Delay Time per Vehicle
ggplot(summary_data, aes(x = Vehicle, y = Avg_Delay_Time, fill = Vehicle)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Delay Time per Vehicle (in Minutes)", x = "Vehicle", y = "Average Delay Time (Minutes)") +
  theme_minimal()

```
Figure 4: Total delay time and average delay time for each type of vehicles

In Figure 4, we observe the trend of total delay time and average delay time for each type of transportation. A closer look at the chart reveals that buses tend to have the highest total delay time and average delay time among all vehicles, with an total delay time of 40,000 minutes and average delay time of approximately 23 minutes which is about 8.8 times longer than the average subway delay time. Eventually, the streetcar has the least total delay time among all the vehicles; however, it has a longer average delay time than the subway.
